### 2020-02-08 doc/icfp20/counterpoint.tex
renaming & text

In this section, we present an implementation of species counterpoint,
based on the formulation given by Fux~\citep{fux-cp}.
The idea is to represent ``good'' counterpoint  as a dependent record,
whose fields are the proofs that the counterpoint follows certain rules.
For space reason, we only describe two variants of species counterpoint;
other variants can be formalized in an analogous way.

@@

We represent such music as a vector of pitch-interval pairs \texttt{(p , i)},
where \texttt{p} is a note in the cantus firmus, and \texttt{i} is the
distance between \texttt{p} and the corresponding counterpoint note.

@@

frog-cp1 : Vector PitchInterval 7
frog-cp1 = (c 5 , per8) :: (d 5 , maj6) :: (e 5 , min6) :: (f 5 , maj3) :: (e 5 , min3) :: (d 5 , maj6) :: (c 5 , per8) :: []
\end{alltt}

Note that the vector representation does not carry any correctness
guarantee of the music;
it requires extrinsic proofs stating that the music follows all the
rules for being good counterpoint.
Below, we discuss five such rules, showing what role they play
and how they can be represented using dependent types.

@@

\begin{alltt}
data BeginningError : Set where
  not158   : PitchInterval \(\rightarrow\) BeginningError
  tooShort : BeginningError

beginningCheck : PitchInterval \(\rightarrow\) Maybe BeginningError
beginningCheck pi@(_ , i) =
  if ((i == per1) \(\vee\) (i == per5) \(\vee\) (i == per8))
  then nothing
  else just (not158 pi)

checkBeginning : List PitchInterval \(\rightarrow\) Maybe BeginningError
checkBeginning []       = just tooShort
checkBeginning (p :: ps) = beginningCheck p  
\end{alltt}

@@

unisonCheck : PitchInterval \(\rightarrow\) Maybe UnisonError
unisonCheck (p , i) = if (i == per1) then just (unison p) else nothing  
\end{alltt}

>>>>

In this section, we present an implementation of species counterpoint,
based on the formulation given by Fux~\citep{fux-cp}.
The idea is to represent ``good'' counterpoint  as a dependent record,
whose fields encode the musical content as well as proofs that the
counterpoint follows certain rules.
For space reason, we only describe two variants of species counterpoint;
other variants can be formalized in an analogous way.

@@

We represent such music in the following way: we keep the first and
last bars as separate elements, and encode the middle bars as a vector
of pitch-interval pairs \texttt{(p , i)}, where \texttt{p} is a note in the
cantus firmus, and \texttt{i} is the distance between \texttt{p} and the
corresponding counterpoint note.
For instance, the main body of the above example is represented as
follows:

@@

cfcp1 : List PitchInterval
cfcp1 = (d 5 , maj6) :: (e 5 , min6) :: (f 5 , maj3) :: (e 5 , min3) :: (d 5 , maj6) :: []
\end{alltt}

The reason we separate different parts of the music is that they are
subject to different rules.
In the reminder of this section, we elaborate these rules one by one.

@@

\begin{alltt}
data BeginningError : Set where
  not158   : PitchInterval \(\rightarrow\) BeginningError

checkBeginning : PitchInterval \(\rightarrow\) Maybe BeginningError
checkBeginning pi@(_ , i) =
  if ((i == per1) \(\vee\) (i == per5) \(\vee\) (i == per8))
  then nothing
  else just (not158 pi)
\end{alltt}

@@

unisonCheck : PitchInterval \(\rightarrow\) Maybe UnisonError
unisonCheck (p , i) = if (i == per1) then just (unison p) else nothing

checkUnison : List PitchInterval \(\rightarrow\) List UnisonError
checkUnison = mapMaybe unisonCheck
\end{alltt}
